Surrogate model for an aligned-spin effective one body waveform model of binary neutron star inspirals using Gaussian process regression

### 2018-12-02 paper/paper_no_event.tex
Paper: corrections in the Appendix.

\subsection{Spin-induced quadrupole-monopole terms}

@@

The authors sincerely thank Zoheyr Doctor, Bhooshan Gadre, and Prayush Kumar for reviewing and testing the implementation of \texttt{SEOBNRv4T\_surrogate} in \texttt{LAL}. BL also thanks the participants of the Bayesian Methods in Nuclear Physics workshop at the University of Washington for helpful advice at the start of this work.

@@

The explicit expressions for the amplitude and phase of the TaylorF2 waveform are as follows. For the amplitude we use the 1PN correction to the leading order waveform.
\begin{equation}
A_{\rm F2} = \sqrt{\frac{5\pi\eta}{24}} v^{-7/2} \left[ 1 + \left(-\frac{323}{224} + \frac{451\eta}{168} \right)v^{2} \right],
\end{equation}
where $v=(\pi M f)^{1/3}$ is the standard PN parameter. This differs slightly from the decomposition used by \texttt{LAL}~\cite{lal} \red{[explain]}. The phase has the schematic form
\begin{align}
\Phi_{\rm F2} =& -2\pi f t_{0} + \phi_{0} + \frac{\pi}{4} - \frac{3}{128\eta}v^{-5} \left[\Phi_{\rm F2}^{\rm PP}(\eta) \right. \nonumber\\
	& \left. \qquad + \Phi_{\rm F2}^{\rm Spin} (\eta, \chi_1, \chi_2) + \Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)  \right] \,,
\end{align}
where $t_{0}$ and $\phi_{0}$ are constants related to the freedom of choosing the time and phase of coalescence. We separated point particle terms $\Phi_{\rm F2}^{\rm PP}(\eta)$, spin terms $\Phi_{\rm F2}^{\rm Spin}(\eta, \chi_{1}, \chi_{2})$ and tidal terms $\Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)$. We include point particle terms to the highest know order of 3.5PN~\cite{}.

The spin terms can be decomposed into spin-orbit and spin-spin terms (ignoring cubic-in-spin contributions at 3.5PN~\cite{}) as $\Phi_{\rm F2}^{\rm Spin} = \Phi_{\rm F2}^{\rm SO}  + \Phi_{\rm F2}^{\rm SS} $. We include spin-orbit terms up to 3.5PN~\cite{}. For spin-square terms, as the \texttt{SEOBNRv4T} waveform does not include yet 3PN spin-spin effects (see Sec.~\ref{}), we stop at  the leading 2PN in the \texttt{TaylorF2} waveform and ignore the known 3PN terms. The main reason for doing this is because including only common physical effects in both \texttt{SEOBNRv4T} and \texttt{TaylorF2} makes the residual $\Delta\Phi(Mf; \bx)$ smaller and have less variation, making it easier to fit at a given error requirement.

@@

The point particle terms coefficients up to 3.5PN can be found e.g. in Eq.~(3.18) of Ref.~\cite{BuonannoIyerOchsner2009}. The non-zero coefficients read
\begin{align}
	a_{2} &= \frac{55 \nu }{9}+\frac{3715}{756} \nonumber\\
	a_{3} &= -16 \pi \nonumber\\
	a_{4} &= \frac{3085 \nu ^2}{72}+\frac{27145 \nu }{504}+\frac{15293365}{508032} \nonumber\\
	a_{5} &= -\frac{65 \pi  \nu }{9}-\frac{65}{3} \pi  \nu   \ln v +\frac{38645}{252} \pi   \ln v +\frac{38645 \pi }{756} \nonumber\\
	a_{6} &= -\frac{127825 \nu ^3}{1296}+\frac{76055 \nu ^2}{1728}+\frac{2255 \pi ^2 \nu }{12} \nonumber\\ 
	& -\frac{15737765635 \nu }{3048192}-\frac{6848  \ln v }{21}-\frac{640 \pi ^2}{3}-\frac{6848 \gamma_{E} }{21} \nonumber\\ 
	& +\frac{11583231236531}{4694215680}-\frac{13696 \log (2)}{21} \nonumber\\
	a_{7} &= -\frac{74045 \pi  \nu ^2}{756}+\frac{378515 \pi  \nu }{1512}+\frac{77096675 \pi }{254016}
\end{align}

The spin-orbit and spin-spin corrections have the structure
\begin{align}
	\Phi_{\rm F2}^{\rm SO}(\eta, \chi_{1}, \chi_{2}) &= \sum_{k=3}^{7} b_{k} v^{k} + \mathcal{O}(v^{8})\nonumber\\
	\Phi_{\rm F2}^{\rm SS}(\eta, \chi_{1}, \chi_{2}) &= c_{4} v^{4} + c_{6} v^{6} + \mathcal{O}(v^{7}) \,.
\end{align}
\begin{widetext}
The non-zero spin-orbit coefficients are
\begin{align}
	b_{3} &= \frac{113 \delta   \chi_a }{3}+\left(\frac{113}{3}-\frac{76 \nu }{3}\right)  \chi_s \nonumber\\
	b_{5} &=  \chi_s  \left(\frac{340 \nu ^2}{9}+\frac{24260 \nu }{81}+\frac{340}{3} \nu ^2  \ln v +\frac{24260}{27} \nu   \ln v -\frac{732985  \ln v }{756}-\frac{732985}{2268}\right) \nonumber\\
	& \quad + \chi_a  \left(-\frac{140 \delta  \nu }{9}-\frac{732985 \delta }{2268}-\frac{140}{3} \delta  \nu   \ln v -\frac{732985}{756} \delta   \ln v \right)  \nonumber\\
	b_{6} &= \frac{2270 \pi  \delta   \chi_a }{3}+\left(\frac{2270 \pi }{3}-520 \pi  \nu \right)  \chi_s \nonumber\\
	b_{7} &=  \chi_a  \left(-\frac{1985 \delta  \nu ^2}{48}+\frac{26804935 \delta  \nu }{6048}-\frac{25150083775 \delta }{3048192}\right)+\left(\frac{5345 \nu ^3}{36}-\frac{1042165 \nu ^2}{3024}+\frac{10566655595 \nu }{762048}-\frac{25150083775}{3048192}\right)  \chi_s \,,
\end{align}
while the leading-order quadratic-in-spin term reads
\begin{align}
	c_{4} &= \chi_s ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \nu -50  \kappa_s -\frac{195 \nu }{2}-\frac{5}{8}\right) + \chi_a   \chi_s  \left(-100 \delta   \kappa_s -\frac{5 \delta }{4}+200  \kappa_a  \nu -100  \kappa_a \right) \nonumber\\
	& + \chi_a ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \nu -50  \kappa_s +100 \nu -\frac{5}{8}\right) \,.
\end{align}
As explained above, to be consistent with the fact that this PN information has not yet been incorporated in \texttt{SEOBNRv4T}, we do not include the next-to-leading 3PN term, which for reference is given by
\begin{align}
	c_{6} &=  \chi_s ^2 \left(-\frac{1495 \delta   \kappa_a  \nu }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \nu ^2-\frac{44255  \kappa_s  \nu }{21}+\frac{26015  \kappa_s }{28}+\frac{3415 \nu ^2}{9}+\frac{829705 \nu }{504}-\frac{1344475}{2016}\right) \nonumber\\
	& + \chi_a   \chi_s  \left(-\frac{1495 \delta   \kappa_s  \nu }{3}+\frac{26015 \delta   \kappa_s }{14}+\frac{745 \delta  \nu }{18}-\frac{1344475 \delta }{1008}-480  \kappa_a  \nu ^2-\frac{88510  \kappa_a  \nu }{21}+\frac{26015  \kappa_a }{14}\right) \nonumber\\
	& + \chi_a ^2 \left(-\frac{1495 \delta   \kappa_a  \nu }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \nu ^2-\frac{44255  \kappa_s  \nu }{21}+\frac{26015  \kappa_s }{28}-240 \nu ^2+\frac{267815 \nu }{252}-\frac{1344475}{2016}\right)\,.
\end{align}
\end{widetext}

Finally, the tidal contributions to the phasing are given by~\cite{VinesFlanaganHinderer2011}
\begin{equation}
	\Phi_{\rm Tidal}^{\rm SS}(\eta, \Lambda_{1}, \Lambda_{2}) = d_{10} v^{10} + d_{12} v^{12} + \mathcal{O}(v^{13})
\end{equation}
Introducing convenient mass-weighted combinations of the tidal parameters, expressed using $X_{1} = m_{1}/(m_{1} + m_{2})$ and $X_{2} = m_{2}/(m_{1} + m_{2})$, \red{[to be checked]}
\begin{align}
	\tilde\Lambda &= \frac{16}{13} \frac{(X_1+12X_2)X_1^4\Lambda_1 + (12X_1+X_2)X_2^4\Lambda_2}{(X_1+X_2)^5} \nonumber\\
	\delta\tilde\Lambda =& \frac{(1319X_1^2 - 7996X_1X_2 - 11005X_2^2)X_1^4\Lambda_1}{1319(X_1+X_2)^6} \\
& + \frac{(11005X_1^2 + 7996X_1X_2 - 1319X_2^2)X_2^4\Lambda_2}{1319(X_1+X_2)^6} \,,
\end{align}
the coefficients are given by \red{[to be checked]}
\begin{align}
	d_{10} &= -\frac{39}{2}\tilde\Lambda \nonumber\\
	d_{12} &= -\frac{3115}{64}\tilde\Lambda + \frac{6595}{364}(X_1-X_2)\delta\tilde\Lambda \,.
\end{align}

>>>>

\subsection{Spin-induced quadrupole-monopole terms}
\label{subsec:QM}

@@

The authors sincerely thank Justin Vines and Tanja Hinderer for deriving the quadrupole-monopole corrections needed in the model. The authors are grateful to Zoheyr Doctor, Bhooshan Gadre, and Prayush Kumar for reviewing and testing the implementation of \texttt{SEOBNRv4T\_surrogate} in \texttt{LAL}. BL also thanks the participants of the Bayesian Methods in Nuclear Physics workshop at the University of Washington for helpful advice at the start of this work.

@@

The explicit expressions for the amplitude and phase of the TaylorF2 waveform are as follows. For the amplitude we use the 1PN correction to the leading order waveform,
\begin{equation}
A_{\rm F2} = \sqrt{\frac{5\pi\eta}{24}} v^{-7/2} \left[ 1 + \left(-\frac{323}{224} + \frac{451\eta}{168} \right)v^{2} \right],
\end{equation}
where $v=(\pi M f)^{1/3}$ is the standard PN parameter. The phase has the schematic form
\begin{align}
\Phi_{\rm F2} =& -2\pi f t_{0} + \phi_{0} + \frac{\pi}{4} - \frac{3}{128\eta}v^{-5} \left[\Phi_{\rm F2}^{\rm PP}(\eta) \right. \nonumber\\
	& \left. \qquad + \Phi_{\rm F2}^{\rm Spin} (\eta, \chi_1, \chi_2) + \Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)  \right] \,,
\end{align}
where $t_{0}$ and $\phi_{0}$ are constants related to the freedom of choosing the time and phase of coalescence. We separated point particle terms $\Phi_{\rm F2}^{\rm PP}(\eta)$, spin terms $\Phi_{\rm F2}^{\rm Spin}(\eta, \chi_{1}, \chi_{2})$ and tidal terms $\Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)$. We include point particle terms to the highest know order of 3.5PN (see e.g.~\cite{Blanchet:2013haa}).

The spin terms can be decomposed into spin-orbit and spin-spin terms (ignoring cubic-in-spin contributions at 3.5PN) as $\Phi_{\rm F2}^{\rm Spin} = \Phi_{\rm F2}^{\rm SO}  + \Phi_{\rm F2}^{\rm SS} $. We include spin-orbit terms up to 3.5PN~\cite{Blanchet:2013haa}. For spin-square terms, as the \texttt{SEOBNRv4T} waveform does not include yet 3PN spin-spin effects (see Sec.~\ref{subsec:QM}), we stop at  the leading 2PN in the \texttt{TaylorF2} waveform and ignore the known 3PN terms. The main reason for doing this is because including only common physical effects in both \texttt{SEOBNRv4T} and \texttt{TaylorF2} makes the residual $\Delta\Phi(Mf; \bx)$ smaller and have less variation, making it easier to fit at a given error requirement.

@@

The point particle terms coefficients up to 3.5PN can be found e.g. in Eq.~(3.18) of Ref.~\cite{BuonannoIyerOchsner2009}. The non-zero coefficients read
\begin{align}
	a_{2} &= \frac{55 \eta }{9}+\frac{3715}{756} \nonumber\\
	a_{3} &= -16 \pi \nonumber\\
	a_{4} &= \frac{3085 \eta ^2}{72}+\frac{27145 \eta }{504}+\frac{15293365}{508032} \nonumber\\
	a_{5} &= -\frac{65 \pi  \eta }{9}-\frac{65}{3} \pi  \eta   \ln v +\frac{38645}{252} \pi   \ln v +\frac{38645 \pi }{756} \nonumber\\
	a_{6} &= -\frac{127825 \eta ^3}{1296}+\frac{76055 \eta ^2}{1728}+\frac{2255 \pi ^2 \eta }{12} \nonumber\\ 
	& -\frac{15737765635 \eta }{3048192}-\frac{6848  \ln v }{21}-\frac{640 \pi ^2}{3}-\frac{6848 \gamma_{E} }{21} \nonumber\\ 
	& +\frac{11583231236531}{4694215680}-\frac{13696 \log (2)}{21} \nonumber\\
	a_{7} &= -\frac{74045 \pi  \eta ^2}{756}+\frac{378515 \pi  \eta }{1512}+\frac{77096675 \pi }{254016}
\end{align}

The spin-orbit and spin-spin corrections have the structure
\begin{align}
	\Phi_{\rm F2}^{\rm SO}(\eta, \chi_{1}, \chi_{2}) &= \sum_{k=3}^{7} b_{k} v^{k} + \mathcal{O}(v^{8})\nonumber\\
	\Phi_{\rm F2}^{\rm SS}(\eta, \chi_{1}, \chi_{2}) &= c_{4} v^{4} + c_{6} v^{6} + \mathcal{O}(v^{7}) \,.
\end{align}
\begin{widetext}
The non-zero spin-orbit coefficients are
\begin{align}
	b_{3} &= \frac{113 \delta   \chi_a }{3}+\left(\frac{113}{3}-\frac{76 \eta }{3}\right)  \chi_s \nonumber\\
	b_{5} &=  \chi_s  \left(\frac{340 \eta ^2}{9}+\frac{24260 \eta }{81}+\frac{340}{3} \eta ^2  \ln v +\frac{24260}{27} \eta   \ln v -\frac{732985  \ln v }{756}-\frac{732985}{2268}\right) \nonumber\\
	& \quad + \chi_a  \left(-\frac{140 \delta  \eta }{9}-\frac{732985 \delta }{2268}-\frac{140}{3} \delta  \eta   \ln v -\frac{732985}{756} \delta   \ln v \right)  \nonumber\\
	b_{6} &= \frac{2270 \pi  \delta   \chi_a }{3}+\left(\frac{2270 \pi }{3}-520 \pi  \eta \right)  \chi_s \nonumber\\
	b_{7} &=  \chi_a  \left(-\frac{1985 \delta  \eta ^2}{48}+\frac{26804935 \delta  \eta }{6048}-\frac{25150083775 \delta }{3048192}\right)+\left(\frac{5345 \eta ^3}{36}-\frac{1042165 \eta ^2}{3024}+\frac{10566655595 \eta }{762048}-\frac{25150083775}{3048192}\right)  \chi_s \,,
\end{align}
while the leading-order quadratic-in-spin term reads
\begin{align}
	c_{4} &= \chi_s ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \eta -50  \kappa_s -\frac{195 \eta }{2}-\frac{5}{8}\right) + \chi_a   \chi_s  \left(-100 \delta   \kappa_s -\frac{5 \delta }{4}+200  \kappa_a  \eta -100  \kappa_a \right) \nonumber\\
	& + \chi_a ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \eta -50  \kappa_s +100 \eta -\frac{5}{8}\right) \,.
\end{align}
As explained above, to be consistent with the fact that this PN information has not yet been incorporated in \texttt{SEOBNRv4T}, we do not include the next-to-leading 3PN term, which for reference is given by~\cite{Bohe:2015ana}
\begin{align}
	c_{6} &=  \chi_s ^2 \left(-\frac{1495 \delta   \kappa_a  \eta }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \eta ^2-\frac{44255  \kappa_s  \eta }{21}+\frac{26015  \kappa_s }{28}+\frac{3415 \eta ^2}{9}+\frac{829705 \eta }{504}-\frac{1344475}{2016}\right) \nonumber\\
	& + \chi_a   \chi_s  \left(-\frac{1495 \delta   \kappa_s  \eta }{3}+\frac{26015 \delta   \kappa_s }{14}+\frac{745 \delta  \eta }{18}-\frac{1344475 \delta }{1008}-480  \kappa_a  \eta ^2-\frac{88510  \kappa_a  \eta }{21}+\frac{26015  \kappa_a }{14}\right) \nonumber\\
	& + \chi_a ^2 \left(-\frac{1495 \delta   \kappa_a  \eta }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \eta ^2-\frac{44255  \kappa_s  \eta }{21}+\frac{26015  \kappa_s }{28}-240 \eta ^2+\frac{267815 \eta }{252}-\frac{1344475}{2016}\right)\,.
\end{align}
\end{widetext}

Finally, the tidal contributions to the phasing take the form~\cite{VinesFlanaganHinderer2011}
\begin{equation}
	\Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_{1}, \Lambda_{2}) = d_{10} v^{10} + d_{12} v^{12} + \mathcal{O}(v^{13}) \,.
\end{equation}
Introducing convenient mass-weighted combinations of the tidal parameters, expressed using $X_{1} = m_{1}/(m_{1} + m_{2})$ and $X_{2} = m_{2}/(m_{1} + m_{2})$,
\begin{align}
	\tilde\Lambda &= \frac{16}{13} \left[(12 - 11 X_1)X_1^4\Lambda_1 + (12 - 11 X_2)X_2^4\Lambda_2 \right] \nonumber\\
	\delta\tilde\Lambda =& \frac{1}{1319} \left[(-11005 + 14014 X_1 - 1690 X_{1}^{2})X_1^4\Lambda_1 \right. \nonumber\\ 
	& \left. \quad\quad + (11005 - 14014 X_2 + 1690 X_{2}^{2}) X_2^4\Lambda_2 \right] \,,
\end{align}
the coefficients are given by
\begin{align}
	d_{10} &= -\frac{39}{2}\tilde\Lambda \nonumber\\
	d_{12} &= -\frac{3115}{64}\tilde\Lambda + \frac{6595}{364}(X_1-X_2)\delta\tilde\Lambda \,.
\end{align}
